---
title: "simpleRegress"
author: "Jason Leon"
date: "Summer 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

The bears neck is 20 inches around which helps predicts about 57 in length.Looking at the accompanying plot, it show that the residual standard error is 5.376, and the R squared is .7524.


## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 


```{r}
MYCHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(MYCHEST)
summary.lm(MYCHEST)
```


The residual standard error is 4.939 and its lower than the previous model which is 5.376and is better. In addition, the R-squared is 0.791, which is higher than the previous model which is 0.7524 and also is better. so in ranking , chest is above in neck sw we can conclude that chest is better than neck.



```{r}
MYHEADWTH <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(MYHEADWTH)
summary.lm(MYHEADWTH)
```


As the third model, the residual standard error is 7.32, which is making it worse. In addition, the R-square is now 0.5409 which is also a decline from the previous model. comparing all three models, this one is at the bottom of the list.


```{r}
MYWEIGHT <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(MYWEIGHT)
summary.lm(MYWEIGHT)
```

This model is a comparison to the bear weight. The residual standard error is 5.432, and the adjusted R-square is 0.7472. In this case it is higher that the last model which is worse, and the R-square is higher than the previous model which is better. So whereas, the residual standard error is worse, the adjusted R-square is better. At this doing the bear's chest has the highest adjusted R-square, which is the best. As far as the standard residual error.


```{r}
MYWEIGHT <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(MYWEIGHT)
summary.lm(MYWEIGHT)
```



This model is a comparison to the bears weights. The residual sandard error is 5.2432 and the adjusted R-squared 0.7423

```{r}
HEADLEN <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(HEADLEN)
summary.lm(HEADLEN)
```

This model is a comparison to bears weight. The residual standards error is 4.235 and the adijusted R-squared 0.8434



